The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-08-07 script/en/chapters/primes.tex
fixed overfull hbox

Due to the fact that the exponents of the Mersenne numbers
\index{Mersenne!number} do not use all
natural numbers, but only the primes, the {\em experimental space} is limited
considerably. The currently known Mersenne prime numbers 
\index{Mersenne!prime number} have the exponents\footnote{%
The following page from Landon Curt Noll\index{Noll, Landon Curt} contains 
all Mersenne primes including its date of discovery and its value as number 
and as word:
      \href{http://www.isthe.com/chongo/tech/math/prime/mersenne.html}
   {\texttt{http://www.isthe.com/chongo/tech/math/prime/mersenne.html}}. \\
Also see:
      \href{http://www.utm.edu/}
   {\texttt{http://www.utm.edu/}}.
                             } 
$$
\begin{array}{c}
2; ~ 3; ~ 5; ~ 7; ~ 13; ~ 17; ~ 19; ~ 31; ~ 61; ~ 89; ~ 107; ~ 127;
~ 521; ~ 607; ~ 1,279; ~ 2,203; ~ 2,281; ~ 3,217; ~ 4,253;\\
 4,423; ~ 9,689; ~ 9,941, ~ 11,213; ~ 19,937; ~ 21,701; ~ 23,207; ~ 44,497; ~
86,243; ~ 110,503; ~ 132,049; \\
 216,091; ~ 756,839; ~ 859,433; ~ 1,257,787; ~ 1,398,269; ~ 2,976,221;
 ~ 3,021,377; ~ 6,972,593; \\
 ~ 13,466,917; ~ 20,996,011; ~ 24,036,583; ~ 25,964,951; ~ 30,402,457; ~ 32,582,657;\\
 ~ 37,156,667; ~ 43,112,609; ~ 42,643,801.
% be_2005_UPDATEN_if-new-mersenne-prime-appears          ~ xxx,xxx,xxx.
\end{array}
$$
Thus    % Eyecatcher_neue-Mersenne
$47$            % be_2005_UPDATEN_if-new-mersenne-prime-appears
Mersenne prime numbers are currently known%
\index{Prime number!Mersenne}\index{Mersenne!prime number}.

@@

% --------------------------------------------------------------------------
\vskip +10 pt
\subsection[Catalan's conjecture]{Catalan's conjecture\footnotemark}
    \footnotetext{%
Eugene Charles Catalan, Belgian mathematician, May 5, 1814$-$Feb 14, 1894.\\
After him the so-called {\em Catalan numbers}
$A(n) = (1 / (n+1) ) * (2n)! / (n!)^2 $ \\
$=  1, 2, 5, 14, 42, 132, 429, 1430, 
4862, 16796, 58786, 208012, 742900, 2674440, 9694845, ... $
are named.
    }
Catalan \index{Catalan Eugene}  \index{Number!Catalan}
conjectured that $ C_4 \;$ is a prime:
$$
 \begin{array}{l}
C_0 = 2, \\
C_1 = 2^{C_0} - 1,  \\
C_2 = 2^{C_1} - 1,  \\
C_3 = 2^{C_2} - 1, \\
C_4 = 2^{C_3} - 1, \cdots \\
\end{array}
$$
    (see {\href{http://www.utm.edu/research/primes/mersenne.shtml}{\tt
http://www.utm.edu/research/primes/mersenne.shtml}} under Conjectures and
Unsolved Problems).

@@

     \begin{table}[h]
     %\begin{center}
     \hspace*{1cm}\begin{tabular}{|r|r|r|r|r|}
     \hline
     {\#} of elements & First element & Distance & When & Discovered by \\ \hline
     22 &  11,410,337,850,553 &   4,609,098,694,200 & 1993 & Paul A. Pritchard,\\
        &                     &                     &      & Andrew Moran,\\
        &                     &                     &      & Anthony Thyssen\\
        &                     &                     &      &\\

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|l|}\hline
Index   &   Precise value  & Rounded value & Comment \\
\hline \hline
1       &   2             &   2  & \\
2       &   3             &   3  &  \\
3       &   5             &   5  & \\
4       &   7             &   7  & \\
5       &   11            &   11 & \\
6       &   13            &   13 & \\
7       &   17            &   17 & \\
8       &   19            &   19 & \\
9       &   23            &   23 & \\
10      &   29            &   29 & \\
100     &   541           &   541 & \\
1,000   &   7,917         &   7,917 & \\
664,559 &   9,999,991     &   9.99999E+06 &  All prime numbers up to 1E+07 were known\\
        &                 &                 &  at the beginning of the 20th century.\\
1E+06   &    15,485,863    &      1.54859E+07 & \\
6E+06   &    104,395,301    &    1.04395E+08  & This prime was discovered in 1959.\\
1E+07   &    179,424,673     &    1.79425E+08 & \\
1E+09   &    22,801,763,489  &    2.28018E+10 & \\
1E+12   &    29,996,224,275,833 & 2.99962E+13 & \\ \hline
\end{tabular}
\caption{List of particular $n$-th prime numbers}
\end{center}
\end{table}

>>>>

Due to the fact that the exponents of the Mersenne numbers
\index{Mersenne!number} do not use all
natural numbers, but only the primes, the {\em experimental space} is limited
considerably. The currently known Mersenne prime numbers 
\index{Mersenne!prime number} have the exponents\footnote{%
The following page from Landon Curt Noll\index{Noll, Landon Curt} contains 
all Mersenne primes including its date of discovery and its value as number 
and as word:
      \href{http://www.isthe.com/chongo/tech/math/prime/mersenne.html}
   {\texttt{http://www.isthe.com/chongo/tech/math/prime/mersenne.html}}. \\
Also see:
      \href{http://www.utm.edu/}
   {\texttt{http://www.utm.edu/}}.
                             } 
$$
\begin{array}{c}
2; ~ 3; ~ 5; ~ 7; ~ 13; ~ 17; ~ 19; ~ 31; ~ 61; ~ 89; ~ 107; ~ 127;
~ 521; ~ 607; ~ 1,279; ~ 2,203; ~ 2,281; ~ 3,217;\\
4,253; ~ 4,423; ~ 9,689; ~ 9,941, ~ 11,213; ~ 19,937; ~ 21,701; ~ 23,207; ~ 44,497; ~
86,243; ~ 110,503; \\
132,049; ~ 216,091; ~ 756,839; ~ 859,433; ~ 1,257,787; ~ 1,398,269; ~ 2,976,221; ~ 3,021,377;\\
6,972,593; ~ 13,466,917; ~ 20,996,011; ~ 24,036,583; ~ 25,964,951; ~ 30,402,457;\\
32,582,657; ~ 37,156,667; ~ 43,112,609; ~ 42,643,801.
% be_2005_UPDATEN_if-new-mersenne-prime-appears          ~ xxx,xxx,xxx.
\end{array}
$$
Thus    % Eyecatcher_neue-Mersenne
$47$            % be_2005_UPDATEN_if-new-mersenne-prime-appears
Mersenne prime numbers are currently known%
\index{Prime number!Mersenne}\index{Mersenne!prime number}.

@@

% --------------------------------------------------------------------------
\vskip +10 pt
\subsection[Catalan's conjecture]{Catalan's conjecture\footnotemark}
    \footnotetext{%
Eugene Charles Catalan, Belgian mathematician, May 5, 1814$-$Feb 14, 1894.\\
After him the so-called {\em Catalan numbers}
$A(n) = (1 / (n+1) ) * (2n)! / (n!)^2 $ \\
$=  1, 2, 5, 14, 42, 132, 429, 1430, 
4862, 16796, 58786, 208012, 742900, 2674440, 9694845, ... $
are named.
    }
\begin{sloppypar}
Catalan \index{Catalan Eugene}  \index{Number!Catalan}
conjectured that $ C_4 \;$ is a prime:
$$
 \begin{array}{l}
C_0 = 2, \\
C_1 = 2^{C_0} - 1,  \\
C_2 = 2^{C_1} - 1,  \\
C_3 = 2^{C_2} - 1, \\
C_4 = 2^{C_3} - 1, \cdots \\
\end{array}
$$
(see \url{http://www.utm.edu/research/primes/mersenne.shtml} under Conjectures and
Unsolved Problems).
\end{sloppypar}

@@

     \begin{table}[h]
     %\begin{center}
     \hspace*{1cm}\begin{tabular}{|r|r|r|r|l|}
     \hline
     Elements & First element & Distance & When & Discovered by \\ \hline
     22 &  11,410,337,850,553 &   4,609,098,694,200 & 1993 & Paul A. Pritchard,\\
        &                     &                     &      & Andrew Moran,\\
        &                     &                     &      & Anthony Thyssen\\
        &                     &                     &      &\\

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|l|}\hline
Index   &   Precise value  & Rounded value & Comment \\
\hline \hline
1       &   2             &   2  & \\
2       &   3             &   3  &  \\
3       &   5             &   5  & \\
4       &   7             &   7  & \\
5       &   11            &   11 & \\
6       &   13            &   13 & \\
7       &   17            &   17 & \\
8       &   19            &   19 & \\
9       &   23            &   23 & \\
10      &   29            &   29 & \\
100     &   541           &   541 & \\
1,000   &   7,917         &   7,917 & \\
664,559 &   9,999,991     &   9.99999E+06 &  All prime numbers up to 1E+07 were\\
        &                 &                 &  known at the beginning of the 20th\\
        &                 &                 &  century.\\
1E+06   &    15,485,863    &      1.54859E+07 & \\
6E+06   &    104,395,301    &    1.04395E+08  & This prime was discovered in 1959.\\
1E+07   &    179,424,673     &    1.79425E+08 & \\
1E+09   &    22,801,763,489  &    2.28018E+10 & \\
1E+12   &    29,996,224,275,833 & 2.99962E+13 & \\ \hline
\end{tabular}
\caption{List of particular $n$-th prime numbers}
\end{center}
\end{table}

### 2009-08-07 script/en/script-en.tex
additional cryptoprocedure usage + consistency corrections
fixed overfull hbox

\documentclass[a4paper,11pt,oneside]{book}
\setlength{\fboxrule}{.5mm}
\setlength{\fboxsep}{1.75mm}
\setlength{\footnotesep}{6pt}

@@

\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage{url}
% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können
%\bibliographyunit[\section] % bibunits für das ganze Dokument aktivieren
\defaultbibliography{references}
\defaultbibliographystyle{abbrv}
\usepackage{float}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,hypertexnames=false,debug]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels and hypertexnames=false and place hyperref after float.
%Option debug helps finding problems.

>>>>

\documentclass[a4paper,11pt,oneside]{book}
\overfullrule=5pt % uncomment to mark overfull boxes
\setlength{\fboxrule}{.5mm}
\setlength{\fboxsep}{1.75mm}
\setlength{\footnotesep}{6pt}

@@

\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage[hyphens]{url} % options "hyphens" enables linebreaks in \url{}
% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können
%\bibliographyunit[\section] % bibunits für das ganze Dokument aktivieren
\defaultbibliography{references}
\defaultbibliographystyle{abbrv}
\usepackage{float}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,hypertexnames=false]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels and hypertexnames=false and place hyperref after float.
%Option debug helps finding problems.
